\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}

### 2024-04-03 lin-vs-cen.tex
Pffft
Fixed calculation error pointed out by referee.

  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>1$ to be defined shortly.  Then
  \begin{align}
    w(E_\alpha)\prod_{E_\beta\in\Gamma_\alpha} (1-w(E_\beta))
    & = \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|-1} \notag \\
    & \ge \exp\left(-\ln\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
    & \ge \exp\left(-\ln\tau - \vol{7r}\cdot|\Phi_\alpha|/(\tau-1)\right)
    & \text{(since $|\Gamma_\alpha|\le \vol{7r}\cdot\Phi_\alpha$).}  \label{weight_product}
  \end{align}

  Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
  Taking $\tau = {32\vol{7r}^2}+1$, and using \cref{weight_product,probability}, we find that
  \[
     \Pr\left(E_\alpha\right) \le w(E_\alpha)\prod_{\beta\in\Gamma_\alpha}(1-w(E_\beta))
  \]
  provided that
  \[   |\Phi_\alpha| \ge {32\vol{7r}}\big(\ln({32\vol{7r}}+1) + {2\vol{2r+1}}\ln(2|\Phi_\alpha|)\big) \enspace .
  \]
  The right hand side of this last equation is of the form $O(r^2\log r+r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d-\vol{r+1}\in d- O(r^2)$, this can therefore be satisfied for some $d\in O(r^4\log r)$.

@@

The main subject of this paper was posed as an open problem at the 9th Annual Bellairs Workshop on Geometry and Graphs (v2), which was held January 21--28, 2022 and again at the 16th Annual Workshop on Probability and Combinatorics, which was held March 25--April 1, 2022. We are grateful to the organizers and participants of both workshops for providing a stimulating research environment.
Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace. We are also grateful to Omer Angel for pointing out that the problem on $k\times k$ grid (as opposed to pseudogrid) has a simple solution.

>>>>

  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>1$ to be defined shortly.  Then
  \begin{align}
    w(E_\alpha)\prod_{E_\beta\in\Gamma_\alpha} (1-w(E_\beta))
    & = \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|-1} \notag \\
    & > \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|}
      & \text{(since $1-1/\tau<1$)}\notag \\
    % & = \exp(-\ln\tau)\cdot \left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|} \notag \\
    % & = \exp(-\ln\tau)\cdot \left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|} \notag \\
    & > \exp(-\ln\tau-|\Gamma_\alpha|/\tau)
      & \text{(since $e^{-1/\tau}> 1-1/\tau$)}\notag \\
    % & \ge \exp\left(-\ln\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
    & \ge \exp\left(-\ln\tau - \vol{7r}\cdot|\Phi_\alpha|/\tau\right)
      & \text{(since $|\Gamma_\alpha|\le \vol{7r}\cdot\Phi_\alpha$).}  \label{weight_product}
  \end{align}

  Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
  Taking $\tau = {32\vol{7r}^2}$, and using \cref{weight_product,probability}, we find that
  \[
     \Pr\left(E_\alpha\right) \le w(E_\alpha)\prod_{\beta\in\Gamma_\alpha}(1-w(E_\beta))
  \]
  provided that
  \[   |\Phi_\alpha| \ge {32\vol{7r}}\big(\ln({32\vol{7r}}+1) + {2\vol{2r+1}}\ln(2|\Phi_\alpha|)\big) \enspace .
  \]
  The right hand side of this last equation is of the form $O(r^2\log r+r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d-\vol{r+1}\in d- O(r^2)$, this can therefore be satisfied for some $d\in O(r^4\log r)$.

@@

The main subject of this paper was posed as an open problem at the 9th Annual Bellairs Workshop on Geometry and Graphs (v2), which was held January 21--28, 2022 and again at the 16th Annual Workshop on Probability and Combinatorics, which was held March 25--April 1, 2022. We are grateful to the organizers and participants of both workshops for providing a stimulating research environment.
Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace. We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution.
